Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-15 resex.tex
Intermediate step in revision of results on random pulsing.

% abbreviations
\newcommand{\seventhtp}{7th-turn pulsing}
\newcommand{\eighthtp}{8th-turn pulsing}
\newcommand{\tenthtp}{10th-turn pulsing}

\newcommand{\fma}{frequency-map analysis}

@@

\item \textbf{random excitation:} The extraction voltage in the
  electron gun is modulated according to the
  following expression:
%
  \begin{equation}
    U_{\mathrm{e-gun}} = (1-a) \cdot U_{\mathrm{max}} +
    a \cdot \eta \cdot U_{\mathrm{max}},
  \end{equation}
%
  where $U_{\mathrm{max}}$ is the maximum voltage, $a$ is the
  modulation strength, with $a \in [0,1]$, and $\eta$ is a uniformly
  distributed random number in the interval~$[0,1]$. Simulations and
  experiments, discussed below, were usually conducted with $a = 1$.

@@

The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL. On the other hand this pattern is also the most dangerous one for the proton beam core. The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the \seventhtp\ as for \eighthtp\ basically no effect was observed. The random excitation is in general so powerful as it excites all frequencies of the beam equally.

This can be seen clearly on the example of the FMA analysis in
amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}. In the case
of a random excitation, the diffusion in tune is increased in
comparison to the $k^{\mathrm{th}}$~turn pulsing rather equally for
all particle amplitudes.

The distribution tracking shown in Fig.~\ref{fig:ransim} also reveals a quite symmetric evolution of the beam parameters for a random excitation:
\begin{itemize}
	\item Emittance growth is only caused by an excitation in the same plane.
	\item The emittance growth is quantitatively comparable for an excitation in only the corresponding plane. Only a small increase or decrease is observed if the excitation is applied in both planes. The difference could be due to coupling.
	\item The random excitation leads to an increase of the emittance growth rate. Compared to the $k^{\mathrm{th}}$~turn pulsing there is explicitly no initial adjustment phase of the distribution.
	\item Negligible losses are observed independent of the plane of excitation.
        \end{itemize}

>>>>

% abbreviations
\newcommand{\kthtp}{$k$th-turn pulsing}
\newcommand{\seventhtp}{7th-turn pulsing}
\newcommand{\eighthtp}{8th-turn pulsing}
\newcommand{\tenthtp}{10th-turn pulsing}

\newcommand{\Fma}{Frequency-map analysis}
\newcommand{\fma}{frequency-map analysis}

@@

\item \textbf{random excitation:} The extraction voltage in the
  electron gun is modulated according to the
  following expression:
%
  \begin{equation}
    \label{eq:random}
    U_{\mathrm{e-gun}} = (1-a) \cdot U_{\mathrm{max}} +
    a \cdot \eta \cdot U_{\mathrm{max}},
  \end{equation}
%
  where $U_{\mathrm{max}}$ is the maximum voltage, $a$ is the
  modulation strength, with $a \in [0,1]$, and $\eta$ is a uniformly
  distributed random number in the interval~$[0,1]$. Simulations and
  experiments, discussed below, were usually conducted with $a = 1$.

@@

The random excitation is a powerful and flexible halo-cleaning
pattern, but it is also potentially harmful for the proton beam
core. A random pattern can be employed at full strength or added as a
modulation to a continuous mode of operation (depending on the
parameter $a$ in Eq.~\ref{eq:random}). During the 2017 experiment, we
tested the effects of the random excitation at full strength ($a = 1$)
and compared it with the resonant excitations.

The random excitation excites practically all resonant frequencies in
the beam.  This can be clearly seen in the FMA example in amplitude
space illustrated in Fig.~\ref{fig:ran2017fmaamp}. Tune diffusion is
increased rather uniformly at all particle amplitudes, and not on
specific resonant lines, as it happens for \kthtp.

The calculated results of distribution tracking, with excitation
amplitudes up to 24~nrad, are shown in
Fig.~\ref{fig:ransim}. They show the following features:
%
\begin{itemize}
\item negligible losses are observed, independent of the plane of
  excitation;
\item there is no significant change in bunch length;
\item emittance growth is only generated by excitations in the same
  plane (or in both planes);
\item horizontal and vertical emittance growth rates are large and
  comparable in magnitude;
\item when the excitation is applied in both planes, a small increase
  or decrease in the emittance growth rate is observed with respect to
  the cases of separate excitations, an effect likely due to lattice
  coupling;
\item emittance growth takes place at a constant rate, without the
  initial adjustment phase characteristic of \kthtp.
\end{itemize}
